### 2021-12-14 results/tables/switch1.tex
Updates from Overleaf

\begin{table}[H]
\centering
\hspace*{-1.2cm}
\begin{threeparttable}
\caption{Effects on yields (log) from adopting improved maize varieties (self-reported)}
\label{tab:switch1}
\begin{tabular}{l cccccc}
\hline
\hline
            &Adopters yields&Non-adopters yields&         ATE&          SE&     p-value\\
\hline
\textit{SR1, SR yields}&            &            &            &            &            \\
ATT         &        7.30&        9.07&       -1.76&       0.011&      0.0000\\
%
%
%
ATU         &        9.06&        6.87&        2.19&       0.024&      0.0000\\
%
%
%
\textit{SR1, cropcut yields}&            &            &            &            &            \\
ATT         &        7.57&        7.86&       -0.30&       0.017&      0.0000\\
%
%
%
ATU         &        8.89&        7.11&        1.78&       0.039&      0.0000\\
%
%
%
\textit{SR1, SR yields, full set of controls}&            &            &            &            &            \\
ATT         &        7.28&        7.82&       -0.54&       0.019&      0.0000\\
%
%
%
ATU         &        9.65&        6.85&        2.80&       0.100&      0.0000\\
%
%
%
\textit{SR1, cropcut yields, full set of controls}&            &            &            &            &            \\
ATT         &        7.56&        7.61&      -0.049&       0.019&      0.0087\\
%
%
%
ATU         &        9.24&        7.11&        2.13&       0.085&      0.0000\\
%
%
%
\textit{SR2, SR yields}&            &            &            &            &            \\
ATT         &        7.30&        7.85&       -0.55&       0.020&      0.0000\\
%
%
%
ATU         &        9.63&        6.88&        2.75&       0.101&      0.0000\\
%
%
%
\textit{SR2, cropcut yields}&            &            &            &            &            \\
ATT         &        7.61&        7.62&      -0.013&       0.019&        0.47\\
%
%
%
ATU         &        9.35&        7.14&        2.21&       0.091&    1.6e-130\\
\hline
\hline
\end{tabular}
\begin{tablenotes}
\footnotesize
\item{Note: Full set of controls include: parcel size (in HA), household labor, hired labor, fertilizer costs, other input costs irrigation (dummy), mechanization (dummy), organic fertilizer (dummy). Instruments for the adoption equation include: years of education of hh head, age of hh head, female head, land title, asset index, seed costs. Results for SR2 include full set of controls. SR1=1 if New, 2nd gen or recycled,=0 if Traditional; SR2=1 if New, =0 2nd gen, recycled or traditional}
\end{tablenotes}
\end{threeparttable}
\end{table}