### 2021-12-14 results/tables/traj.tex
Updates from Overleaf

\begin{table}
\centering
\caption{Frequencies of Trajectories of Households}
\label{tbl:Trajectories}
\begin{tabular}{lr}
\toprule
{} &  Trajectory \\
\midrule
000 &        2124 \\
111 &         456 \\
011 &         228 \\
001 &         219 \\
010 &         153 \\
100 &         108 \\
110 &          96 \\
101 &          87 \\
\bottomrule
\end{tabular}
\end{table}