### 2022-01-27 results/tables/traj.tex

\begin{table}
\centering
\caption{Frequencies of Trajectories of Households}
\label{tbl:Trajectories}
\begin{tabular}{lcc}
\toprule
{} &  Trajectory & Share \\
\midrule
000 &        2,124 &  0.61  \\
111 &         456 &  0.13  \\
011 &         228 &  0.07  \\
001 &         219 &  0.06  \\
010 &         153 &  0.04  \\
100 &         108 &  0.03  \\
110 &          96 &  0.03  \\
101 &          87 &  0.03  \\
\bottomrule
\end{tabular}
\vspace*{-2em}
\begin{table}[H]
    \centering
    \begin{tabular}{p{0.5\textwidth}} 
        \begin{tablenotes}
              \small
              \item Note: Author's calculations using ESS.
        \end{tablenotes}
    \end{tabular}
\end{table}    
\end{table}