### 2022-07-22 results/tables/misclass_95purity.tex
add misclassification matrix

\begin{table}
\caption{Misclassification Matrix for Seed Adoption based on DNA Fingerprinting}
\label{tab:cm}
\begin{tabular}{lllllllll}
\toprule
 & 000 & 001 & 010 & 011 & 100 & 101 & 110 & 111 \\
\midrule
000 & 0.642736 & 0.102033 & 0.102033 & 0.016198 & 0.102033 & 0.016198 & 0.016198 & 0.002571 \\
001 & 0.220303 & 0.524466 & 0.034973 & 0.083258 & 0.034973 & 0.083258 & 0.005552 & 0.013217 \\
010 & 0.220303 & 0.034973 & 0.524466 & 0.083258 & 0.034973 & 0.005552 & 0.083258 & 0.013217 \\
011 & 0.075510 & 0.179765 & 0.179765 & 0.427960 & 0.011987 & 0.028537 & 0.028537 & 0.067938 \\
100 & 0.220303 & 0.034973 & 0.034973 & 0.005552 & 0.524466 & 0.083258 & 0.083258 & 0.013217 \\
101 & 0.075510 & 0.179765 & 0.011987 & 0.028537 & 0.179765 & 0.427960 & 0.028537 & 0.067938 \\
110 & 0.075510 & 0.011987 & 0.179765 & 0.028537 & 0.179765 & 0.028537 & 0.427960 & 0.067938 \\
111 & 0.025882 & 0.061616 & 0.061616 & 0.146687 & 0.061616 & 0.146687 & 0.146687 & 0.349211 \\
\bottomrule
\end{tabular}
\end{table}