### 2022-11-01 results/tables/misclass_95purity.tex
Updates from Overleaf

\begin{table}
\caption{Misclassification Matrix for Seed Adoption based on DNA Fingerprinting}
\label{tab:cm}
\centering
\begin{tabular}{lllllllll}
\toprule
 & 000 & 001 & 010 & 011 & 100 & 101 & 110 & 111 \\
\midrule
000 & 0.643 & 0.102 & 0.102 & 0.016 & 0.102 & 0.016 & 0.016 & 0.003 \\
001 & 0.220 & 0.524 & 0.035 & 0.083 & 0.035 & 0.083 & 0.006 & 0.013 \\
010 & 0.220 & 0.035 & 0.524 & 0.083 & 0.035 & 0.006 & 0.083 & 0.013 \\
011 & 0.076 & 0.180 & 0.180 & 0.428 & 0.012 & 0.029 & 0.029 & 0.068 \\
100 & 0.220 & 0.035 & 0.035 & 0.006 & 0.524 & 0.083 & 0.083 & 0.013 \\
101 & 0.076 & 0.180 & 0.012 & 0.029 & 0.180 & 0.428 & 0.029 & 0.068 \\
110 & 0.076 & 0.012 & 0.180 & 0.029 & 0.180 & 0.029 & 0.428 & 0.068 \\
111 & 0.026 & 0.062 & 0.062 & 0.147 & 0.062 & 0.147 & 0.147 & 0.349 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
    \item \footnotesize Notes: Table shows results of misclassification probabilities for each trajectory in the sample, using probabilities from Table \ref{tab:a}. Misclassification probabilities calculated from DNA fingerprinting data from Wave 4 of the ESS and assumes that true adoption of hybrid seed is when DNA purity of hybrid is at least 95\%.
\end{tablenotes}
\end{table}