### 2023-03-20 results/tables/traj.tex
change number format for traj table

\begin{table}
\caption{Trajectories of Households}
\label{tbl:trajectories}
\begin{tabular}{lrr}
\toprule
 & Frequency & Share \\
Trajectory &  &  \\
\midrule
000 & 2,124 & 0.612 \\
111 & 456 & 0.131 \\
011 & 228 & 0.066 \\
001 & 219 & 0.063 \\
010 & 153 & 0.044 \\
100 & 108 & 0.031 \\
110 & 96 & 0.028 \\
101 & 87 & 0.025 \\
\bottomrule
\end{tabular}
\end{table}