### 2023-03-20 results/tables/traj.tex
Updates from Overleaf

\begin{table}
\centering
\caption{Trajectories of Households}
\label{tbl:trajectories}
\begin{tabular}{lrr}
\toprule
Trajectory &  Frequency &    Share \\
\midrule
       000 &       2124 & 0.611927 \\
       111 &        456 & 0.131374 \\
       011 &        228 & 0.065687 \\
       001 &        219 & 0.063094 \\
       010 &        153 & 0.044080 \\
       100 &        108 & 0.031115 \\
       110 &         96 & 0.027658 \\
       101 &         87 & 0.025065 \\
\bottomrule
\end{tabular}
\end{table}